---
title: "functions-demo"
output:
  html_document:
    theme: readable
    highlight: zenburn
    toc: true
params: 
  url:
    value: x  
---



<br>
<hr>
<br>

I was confused on the directions for creating the functions demo markdown. I included my two attempts at what was being asked in order to capture my thinking/attempts. This markdown shows a test of the salary-report markdown to ensure the functions work.

### Example Report Assets Section 


```{r setup, include=FALSE}
knitr::opts_chunk$set( echo = TRUE, message=F, warning=F )
```


```
# Be sure to hide all of the extra messages and code in the final report:
# ```{r setup, include=FALSE}
# knitr::opts_chunk$set( echo = F, message=F, warning=F )
# ```
```


```{r libraries}
# LOAD PACKAGES
library( dplyr )
library( pander )
library( knitr )
library( gender )

# SOURCE CUSTOM FUNCTIONS
source( "utils.R" )
```




```{r data}
# LOAD DATA
# URL <- params$url
# d <- read.csv( URL )

# FOR MANUAL TESTING ONLY
URL <- 'https://raw.githubusercontent.com/Watts-College/paf-514-template/main/labs/batch-demo/asu-salaries-2020.csv'
d <- read.csv( URL )

# DATA STEPS 

d$first.name <- get_first_name( d$Full.Name )
d <- add_gender( d )
d <- add_titles( d )
d <- fix_salary( d )

```


<br>
<hr>
<br>





# Salary Range by Rank and Gender 

```{r, results="asis", fig.height=7, fig.width=10}
t.salary <- 
  d %>% 
  filter( ! is.na( title ) & title != "" ) %>% 
  group_by( title, gender ) %>% 
  summarize( q25=quantile(salary,0.25),
             q50=quantile(salary,0.50),
             q75=quantile(salary,0.75),
             n=n() ) %>% 
  ungroup() %>% 
  mutate( p= round( n/sum(n), 2) )

pander( t.salary )

t.salary %>% knitr::kable(format="html")
```



<br>
<hr>
<br>

```{r, fig.height=7, fig.width=10, results="asis"}
for( i in academic.units )
{
  
  d2 <- filter( d, Department.Description == i )
  if( nrow(d2) == 0 ) { next }
  # create the salary summary by rank & gender table 
  # create the top 5 salaries table 
  # build the graph 
  d2 %>%  
  filter( Department.Description == i ) %>% 
  create_salary_table( ) %>% 
  build_graph( unit=i )

  # print the tables:  
 #  cat( paste('<h3>',"TOP FIVE SALARIES",'</h3>' ) )
  cat( t.salary %>% knitr::kable(format="html") )   

  # cat( '<br><hr><br>' )
  
}
```



<style>

td {
    padding: 3px 15px 3px 15px;
    text-align: center !important;
}

th {
    padding: 3px 15px 3px 15px;
    text-align: center !important;
    font-weight: bold;
    color: SlateGray !important;
}

h1 { 
    color: SlateGray !important;
    font-size: 2.05em;
    margin-top: 25px;
    
}

h3 { 
    color: SlateGray;
    margin-top: 25px;
}

</style>




